WIDS 2024 Final Project

### 2024-01-09 Code/R FIle.qmd
effective minimum wage graph

#### Graphing Full State Data
##### For facets, use best judgement if by state, decade, year, or region
##### maybe make all states light grey and just choose outline outlier states and maybe regional averages

##### Party by state and year
##### Min Wage by state and year (include fed (facet or use by decades where necessary)
##### Adjusted Min Wage by state and year (include adjusted fed (facet or use by decades where necessary)
##### Union employment, CB employment, and full employment (facet by decade)
##### percent union and percent collective bargaining

>>>>

#### Graphing Full State Data

##### For facets, use best judgement if by state, decade, year, or region

##### maybe make all states light grey and just choose outline outlier states and maybe regional averages

##### Party by state and year

```{r}
ggplot(full_state_data, aes(x = state_abbreviation, y = )) +
```

##### Min Wage by state and year (include fed (facet or use by decades where necessary)

# figure out spacing of graph

# can't do a line for federal minimum wage unless faceted by year instead of decade

```{r}
effective_min_wage_adj_graph1 <- full_state_data |> 
  filter(!is.na(regions)) |> 
  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
  filter(state_abbreviation != "HI") |> 
  ggplot(aes(x = year)) +
  geom_line(aes(y = effective_min_wage_2020_adj, color = state_abbreviation)) +
  scale_color_manual(values = c(rep("gray", 34))) +
  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75)

HI_effective_min_wage <- full_state_data |> 
  filter(!is.na(regions)) |> 
  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
  filter(state_abbreviation == "HI") |> 
  mutate(effective_min_wage_2020_adj = effective_min_wage_2020_adj - .05)

CT_effective_min_wage <- full_state_data |> 
  filter(!is.na(regions)) |> 
  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
  filter(state_abbreviation == "CT")

AK_effective_min_wage <- full_state_data |> 
  filter(!is.na(regions)) |> 
  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
  filter(state_abbreviation == "AK")

CA_effective_min_wage <- full_state_data |> 
  filter(!is.na(regions)) |> 
  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
  filter(state_abbreviation == "CA")

effective_min_wage_adj_graph1 +
  geom_line(data = HI_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "blue",linewidth = .7) +
  annotate("text", x=1994, y=9.5, label= "Hawaii", color = "blue") +
  geom_line(data = CT_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "#288225", linewidth = .7) +
  annotate("text", x=1985, y=7.5, label= "Connecticut", color = "#288225") +
  geom_line(data = AK_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "orange", linewidth = .7) +
  annotate("text", x=1985, y=10, label= "Alaska", color = "orange") +
  geom_line(data = CA_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "purple", linewidth = .7) +
  annotate("text", x=2016, y=12, label= "California", color = "purple") +
  annotate("text", x=2000, y=5.5, label= "Federal", color = "red") +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "State effective minimum wage by Year (adjusted for 2020 $)",
       subtitle = "For states with an adjusted effective minimum wage above the federal level",
       x = "Year",
       y = "Effective minimum wage (2020 USD)")

rm(HI_effective_min_wage, CT_effective_min_wage,
   AK_effective_min_wage, CA_effective_min_wage)
```

```{r}
# abbreviation and decade
ggplot(full_state_data, aes(x = state_abbreviation)) +
  geom_point(aes(y = state_min_wage)) +
  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
  facet_wrap(~decade, nrow = 5) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(x = "State", 
       y = "State Minimum Wage")

# region and decade
ggplot(full_state_data, aes(x = regions)) +
  geom_point(aes(y = state_min_wage)) +
  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
  facet_wrap(~decade, nrow = 5) +
  labs(x = "State", 
       y = "State Minimum Wage")

# region and year
ggplot(full_state_data, aes(x = regions)) +
  geom_point(aes(y = state_min_wage)) +
  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
  facet_wrap(~year, nrow = 10) +
  labs(x = "State", 
       y = "State Minimum Wage")
```

##### Adjusted Min Wage by state and year (include adjusted fed (facet or use by decades where necessary)

# figure out spacing of graph

# can't do a line for adj federal minimum wage unless faceted by year instead of decade

# add a label for the line saying federal minimum wage

```{r}
# abbreviation and decade
ggplot(full_state_data, aes(x = state_abbreviation)) +
  geom_point(aes(y = state_min_wage_2020_adj)) +
  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
  facet_wrap(~decade, nrow = 5) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(x = "State", 
       y = "State Minimum Wage in 2020 Dollars")

# region and decade
ggplot(full_state_data, aes(x = regions)) +
  geom_point(aes(y = state_min_wage_2020_adj)) +
  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
  facet_wrap(~decade, nrow = 5) +
  labs(x = "Region", 
       y = "State Minimum Wage in 2020 Dollars")

# region and year
ggplot(full_state_data, aes(x = regions)) +
  geom_point(aes(y = state_min_wage_2020_adj)) +
  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
  facet_wrap(~year, nrow = 10) +
  labs(x = "Region", 
       y = "State Minimum Wage in 2020 Dollars")
```

##### Union employment, CB employment, and full employment (facet by decade)

# should we take the average value per decade?

# add a key for the color of the points

# fix spacing because union employment count is not visible

```{r}
# state and decade
ggplot(full_state_data, aes(x = state_abbreviation)) +
  geom_point(aes(y = employment_count), color = "red") +
  geom_point(aes(y = union_employment_count), color = "blue") +
  geom_point(aes(y = collective_bargain_employment_count), color = "green") +
  facet_wrap(~decade, nrow = 5) +
  labs(x = "State", 
       y = "Number of People Employed")

# region and decade
ggplot(full_state_data, aes(x = regions)) +
  geom_point(aes(y = employment_count), color = "red") +
  geom_point(aes(y = union_employment_count), color = "blue") +
  geom_point(aes(y = collective_bargain_employment_count), color = "green") +
  facet_wrap(~decade, nrow = 5) +
  labs(x = "Region", 
       y = "Number of People Employed")
```

##### percent union and percent collective bargaining

# should we take the average value per decade?

# add a key for the color of the points

```{r}
# state and decade
ggplot(full_state_data, aes(x = state_abbreviation)) +
  geom_point(aes(y = percent_union), color = "red") +
  geom_point(aes(y = percent_collective_bargain), color = "blue") +
  facet_wrap(~decade, nrow = 5) +
  labs(x = "State", 
       y = "Number of People Employed")

# region and decade
ggplot(full_state_data, aes(x = regions)) +
  geom_point(aes(y = percent_union), color = "red") +
  geom_point(aes(y = percent_collective_bargain), color = "blue") +
  facet_wrap(~decade, nrow = 5) +
  labs(x = "Region", 
       y = "Number of People Employed")
```
